---
title: "New Program Market Analysis"
author: "Ryan Lewis"
date: "3/30/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


<!---Load libraries--->
```{r load_libraries, warning=FALSE}
library(readxl)
library(dplyr)
library(stringr)
```


<!---Define functions--->
```{r functions}
percent_change <- function(num1, num2)  {
  return((num2-num1)/num1*100)
}

```


Questions to answer:

* What programs or certificates would adult learners be interested in?
+ What degrees/certificates have seen large or increasing numbers of completers
+ What professions contain high numbers of adult learners without degrees?
+ What professions are likely to have adult learners that need certificates to demonstrate continuing education/self-improvement?
+ What degrees or certificates might be interesting to those who are looking for self-fulfillment?

There is an assumption that the programs will be online.

<!---
Table list for IPEDS:
  HD2014: Directory information
  IC2014: Program and award level offerings
  IC2014_AY: Student charges for year
  EF2014A_DIST: Distance education status and level of students
  C2014_A: Awards/degrees conferred by program (6-digit CIP code), award level
  C2014_C: Number of students receiving awards/degrees, by award level
  C2014DEP: Number of programs offered and number of programs offered via distance 
  education
  FLAGS<YYYY>: Student response information
--->



## IPED Data codes and descriptions
IPEDS data is represented in numeric form to conserve space. This section maps the numeric values to their corresponding text values for better readability.

```{R Data Mapping}
# CIP code mapping to descriptions
CIP_mapping <- read_excel("~/westminster_program_research/CIP_mapping.xlsx")

# Institutional Characteristics
institution_level <- c("4 year or more","2 to 4 year","less than 2 year","unkown")
institution_type <- c("Public","Private not-for profit","Private for-profit","not available")
institution_highestlevel <- c("Award of less than one academic year","At least 1, but less than 2 academic yrs","Associate's degree","At least 2, but less than 4 academic yrs","Bachelor's degree","Postbaccalaureate certificate","Master's degree","Post-master's certificate","Doctor's degree","Not applicable, first-professional only","{Not available}")
institution_sector <- c("Administrative Unit","Public, 4-year or above","Private not-for-profit, 4-year or above","Private for-profit, 4-year or above","Public, 2-year","Private not-for-profit, 2-year","Private for-profit, 2-year","Public, less-than 2-year","Private not-for-profit, less-than 2-year","Private for-profit, less-than 2-year","Sector unknown (not active)")
institution_UG <- c("Undergraduate degree or certificate offering","No undergraduate offering","Not available")
institution_GR <- c("Graduate degree or certificate offering","No graduate offering","Not available")
institution_highestGR <- c("Doctor's degree - research/scholarship and professional practice","Doctor's degree - research/scholarship","Doctor's degree -  professional practice","Doctor's degree - other","Master's degree","Bachelor's degree","Associate's degree","Non-degree granting","Not available")
institution_degreegrantingstatus <- c("Degree-granting","Nondegree-granting, primarily postsecondary","Not available")

# Institutions variables needed for analysis
institution_columns <- c("UNITID","INSTNM","ADDR","CITY","STABBR","ZIP","FIPS","OBEREG","WEBADDR","SECTOR","ICLEVEL","CONTROL","HLOFFER","UGOFFER","GROFFER","HDEGOFR1","DEGGRANT","CLOSEDAT")



```

## Target Markets


```{R Target Markets}
# Vector of state(s)
market_state <- c("UT")

# vector of IPED UNITID for home Institution
home_institution <- c(230807)

# Vector of IPED UNITID(s) for peer institution(s)
market_peer_inst <- c(180106,209065)

# Comparision institutions
comparison_inst <-c(home_institution,market_peer_inst)
```



```{R Institutions 2016}
institutions_2016 <- read.csv("~/westminster_program_research/IPEDS/HD2016.txt", header = TRUE)

# Select only needed columns
institutions_2016 <- select(institutions_2016, one_of(institution_columns))

# Factorize key variables

#institutions_2016$CONTROL1 <- recode_factor(institutions_2016$CONTROL, `1` = "Public", `2` = "Private not-for profit", `3` = "Private for-profit", `-3` = "not #available")
institutions_2016$CONTROL <- factor(institutions_2016$CONTROL) %>% plyr::mapvalues(from = c("1","2","3","-3"), to = institution_type, warn_missing = FALSE )
institutions_2016$ICLEVEL <- factor(institutions_2016$ICLEVEL) %>% plyr::mapvalues(from = c("1","2","3","-3"), to = institution_level, warn_missing = FALSE )
institutions_2016$SECTOR <- factor(institutions_2016$SECTOR) %>% plyr::mapvalues(from = c("0","1","2","3","4","5","6","7","8","9","99"), to = institution_sector, warn_missing = FALSE )
institutions_2016$HLOFFER <- factor(institutions_2016$HLOFFER) %>% plyr::mapvalues(from = c("1","2","3","4","5","6","7","8","9","-2","-3"), to = institution_highestlevel, warn_missing = FALSE )
institutions_2016$UGOFFER <- factor(institutions_2016$UGOFFER) %>% plyr::mapvalues(from = c("1","2","-3"), to = institution_UG, warn_missing = FALSE )
institutions_2016$GROFFER <- factor(institutions_2016$GROFFER) %>% plyr::mapvalues(from = c("1","2","-3"), to = institution_GR, warn_missing = FALSE )
institutions_2016$HDEGOFR1 <- factor(institutions_2016$HDEGOFR1) %>% plyr::mapvalues(from = c("11","12","13","14","20","30","40","0","-3"), to = institution_highestGR, warn_missing = FALSE )
institutions_2016$DEGGRANT <- factor(institutions_2016$DEGGRANT) %>% plyr::mapvalues(from = c("1","2","-3"), to = institution_degreegrantingstatus, warn_missing = FALSE )



# Create seperate lists for closed and open institutions. Closed institutuions have a date value in the CLOSEDAT field so searching for a "/" is a good way to distinguish between the two.
closed_institutions_2016 <- institutions_2016 %>% filter(str_detect(CLOSEDAT,"/"))
open_institutions_2016 <- anti_join(institutions_2016,closed_institutions_2016, by = "UNITID")

# For our purposes, we are only interested in degree granting institutions


# Define compartive markets
institutions_2016_homestate <- filter(open_institutions_2016, STABBR %in% market_state) %>% filter(DEGGRANT == "Degree-granting")
institutions_2016_peergroup <- filter(open_institutions_2016, UNITID %in% comparison_inst)



```

## Degree Completion
Information about degree completion was obtain through The Integrated Postsecondary Education Data System (IPEDS) from https://nces.ed.gov/ipeds/use-the-data/download-access-database.

http://bluechipsolutions.com/BI_University/CIP_Codes_Definition_and_Link_to_Download_to_Excel_or_Access.htm

*Questions:
+ What trends exist in degrees/certifictes granted (C2016A)? This will give us an idea of what is growing.
+ What programs are offered and in what format (C2016DEP)?
+ What are local competitors doing in those areas?
+ What does the online/on-campus landscape look like?

<!--- load IPED completion data --->
```{R Program Completion}
#IPEDS_2014 <- read.csv("~/westminster_program_research/IPEDS/C2014_A.txt", header = TRUE)
#IPEDS_2015 <- read.csv("~/westminster_program_research/IPEDS/C2015_A.txt", header = TRUE)

# Set the data types for the IPEDS completion data
temp <- c(rep("integer",34))
temp[2] <- "character"
IPEDS_2016 <- read.csv("~/westminster_program_research/IPEDS/C2016_A.txt", header = TRUE, colClasses = temp)
IPEDS_2016$Year <- "2016"

prog_completion <- IPEDS_2016 %>% select("UNITID","CIPCODE","MAJORNUM","CIPCODE","AWLEVEL","CTOTALT")
#prog_top_completed <- prog_completion %>%
#  group_by(CIPCODE) %>%
#  summarise(total = sum(CTOTALT)) %>%
#  top_n(20, total) %>%
#  arrange(desc(total)) %>%
#  View()

prog_completion_peer <- prog_completion %>% filter(UNITID %in% institutions_2016_peergroup$UNITID)
prog_completion_state <- prog_completion %>% filter(UNITID %in% institutions_2016_homestate$UNITID)

```







```{R Program Offerings 2016}
program_offerings_2016 <- read.csv("~/westminster_program_research/IPEDS/IC2016.txt")

```




```{R Distance Education}
temp <- c(rep("integer",26))
temp[2] <- "character"
onlineprogs_2016 <- read.csv("~/westminster_program_research/IPEDS/C2016DEP.txt", header = TRUE, colClasses = temp)
```


  
<!--- Map IPED and OES data --->

<!--- Define state and local institutions --->







## Labor Market
Using the occupational employment statistic (oes) data from the Bureau of Labor Statistics (https://www.bls.gov/oes/tables.htm), we examine trends in both the Utah labor market and national labor market from 2014 to 2017.

<!--- load state data--->
<!---
```{r echo=FALSE, warning=FALSE}

oes_2014_state <- read_excel("~/market_research/OES/state_M2014_dl.xlsx")
oes_2014_state$YEAR <- "2014"


oes_2015_state <- read_excel("~/market_research/OES/state_M2015_dl.xlsx")
oes_2015_state$YEAR <- "2015"

oes_2016_state <- read_excel("~/market_research/OES/state_M2016_dl.xlsx")
oes_2016_state$YEAR <- "2016"

oes_2017_state <- read_excel("~/market_research/OES/state_M2017_dl.xlsx")
oes_2017_state$YEAR <- "2017"

oes_state <- bind_rows(oes_2014_state,oes_2015_state,oes_2016_state,oes_2017_state)
oes_state$TOT_EMP <- as.numeric(oes_state$TOT_EMP)
oes_utah <- filter(oes_state,STATE=="Utah")

```
--->

<!---load national data--->
<!---
```{r echo=FALSE, warning=FALSE}
oes_2014_national <- read_excel("~/market_research/OES/national_M2014_dl.xlsx")
oes_2014_national$YEAR <- "2014"

oes_2015_national <- read_excel("~/market_research/OES/national_M2015_dl.xlsx")
oes_2015_national$YEAR <- "2015"

oes_2016_national <- read_excel("~/market_research/OES/national_M2016_dl.xlsx")
oes_2016_national$YEAR <- "2016"

oes_2017_national <- read_excel("~/market_research/OES/national_M2017_dl.xlsx")
oes_2017_national$YEAR <- "2017"

oes_national <- bind_rows(oes_2014_national,oes_2015_national,oes_2016_national,oes_2017_national)
oes_national$TOT_EMP <- as.numeric(oes_national$TOT_EMP)
```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
--->
