---
title: "New Program Market Analysis"
author: "Ryan Lewis"
date: "3/30/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


<!---Load libraries--->
```{r warning=FALSE}
library(readxl)
library(dplyr)
library(stringr)
```


<!---Define functions--->
```{r echo=FALSE}
percent_change <- function(num1, num2)  {
  return((num2-num1)/num1*100)
}

```


Questions to answer:

* What programs or certificates would adult learners be interested in?
+ What degrees/certificates have seen large or increasing numbers of completers
+ What professions contain high numbers of adult learners without degrees?
+ What professions are likely to have adult learners that need certificates to demonstrate continuing education/self-improvement?
+ What degrees or certificates might be interesting to those who are looking for self-fulfillment?

There is an assumption that the programs will be online.

## Degree Completion
Information about degree completion was obtain through The Integrated Postsecondary Education Data System (IPEDS) from https://nces.ed.gov/ipeds/use-the-data/download-access-database.

*Questions:
+ What trends exist in degrees/certifictes granted (C2016A)? This will give us an idea of what is growing.
+ What programs are offered and in what format (C2016DEP)?
+ What are local competitors doing in those areas?
+ What does the online/on-campus landscape look like?

<!--- load IPED completion data --->
```{R}
#IPEDS_2014 <= read.csv("~/market_research/IPEDS/C2014_A.txt", header = TRUE)
#IPEDS_2015 <= read.csv("~/market_research/IPEDS/C2015_A.txt", header = TRUE)
IPEDS_2016 <= read.csv("~/market_research/IPEDS/C2016_A.txt", header = TRUE)

```


## Labor Market
Using the occupational employment statistic (oes) data from the Bureau of Labor Statistics (https://www.bls.gov/oes/tables.htm), we examine trends in both the Utah labor market and national labor market from 2014 to 2017.

<!--- load state data--->
```{r echo=FALSE, warning=FALSE}

oes_2014_state <- read_excel("~/market_research/OES/state_M2014_dl.xlsx")
oes_2014_state$YEAR <- "2014"


oes_2015_state <- read_excel("~/market_research/OES/state_M2015_dl.xlsx")
oes_2015_state$YEAR <- "2015"

oes_2016_state <- read_excel("~/market_research/OES/state_M2016_dl.xlsx")
oes_2016_state$YEAR <- "2016"

oes_2017_state <- read_excel("~/market_research/OES/state_M2017_dl.xlsx")
oes_2017_state$YEAR <- "2017"

oes_state <- bind_rows(oes_2014_state,oes_2015_state,oes_2016_state,oes_2017_state)
oes_state$TOT_EMP <- as.numeric(oes_state$TOT_EMP)
oes_utah <- filter(oes_state,STATE=="Utah")

```

<!---load national data--->
```{r echo=FALSE, warning=FALSE}
oes_2014_national <- read_excel("~/market_research/OES/national_M2014_dl.xlsx")
oes_2014_national$YEAR <- "2014"

oes_2015_national <- read_excel("~/market_research/OES/national_M2015_dl.xlsx")
oes_2015_national$YEAR <- "2015"

oes_2016_national <- read_excel("~/market_research/OES/national_M2016_dl.xlsx")
oes_2016_national$YEAR <- "2016"

oes_2017_national <- read_excel("~/market_research/OES/national_M2017_dl.xlsx")
oes_2017_national$YEAR <- "2017"

oes_national <- bind_rows(oes_2014_national,oes_2015_national,oes_2016_national,oes_2017_national)
oes_national$TOT_EMP <- as.numeric(oes_national$TOT_EMP)
```


```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
